---
title: "Lab part II week 9"
author: "Lindsay Hafen"
date: "12/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(broom)
```

## Read in data

```{r}
lizards <- read_csv(here("data", "lizards.csv"))
```

Get counts of lizards by species:
```{r}
lizards %>% 
  count(spp) %>% 
  arrange(-n)

# UTST is the most common, so that is the species we will continue with
```

## Make a subset

- UTST only
- Highest trap sites: IBPE (enclosure grasslands), RABB (Rabbit mesquite site), "EAST" (East tarbrush site)
- Only keep observations where tail condition is recorded (W = whole, B = broken)

```{r}
utst <- lizards %>% 
  filter(spp == "UTST") %>% 
  filter(site %in% c("IBPE", "RABB", "EAST")) %>% #Use unique in console to double check unique(utst$site)
  filter(tail %in% c("B", "W"))
```

## Find counts and proportions of lizards with broken or whole tails at each site

```{r}
utst %>% 
  count(site, tail)

# Cool, but in long format. If you want in short format for chi-squared etc, use janitor's tabyl function (Below)
```

```{r}
tail_counts <- utst %>% 
  janitor::tabyl(site, tail)

# Also show proportions directly!

tail_proportions <- tail_counts %>% 
  janitor::adorn_percentages() %>% 
  janitor::adorn_pct_formatting(digits = 2) %>% #To round percentages to 2 digits
  janitor::adorn_ns() # Shows actual counts as well!!!
```

- Is there enough evidence to reject the H0 that tail breakage is independent of site?
- Is there a significant association between site and tail breakage?
- It looks like there is something going on at the Rabbit site...

## Convert to contingency table
- Use column to rowname function to switch site from data to titles instead of numbers (1,2,3)

```{r}
lizard_ct <- tail_counts %>% 
  column_to_rownames(var = "site")
```

## Run Chi-square
- Are tail breakage and site independent?

```{r}
lizard_x2 <- chisq.test(lizard_ct)
lizard_x2

# Use broom tidy function to display tidy data frame in text reference

lizard_tidy <- broom::tidy(lizard_x2)
```
- Probability that we would get counts at least this different from the expected results under the null hypothesis is .1456%
- We would therefore reject the H0
- Site and lizard tail breakage are NOT independent

Example of inline referencing:

There is a significant association (i.e. non-independence) between site and lizard tail breakage: ($\chi$^2^(`r lizard_tidy$parameter`) = `r lizard_tidy$statistic`, *p* = `r lizard_tidy$p.value`).


